Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election 

### 2024-04-12 paper/paper(1).qmd
Add files via upload

---
title: "Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election "
title-block-banner: true
abstract: "ojjojojojljk,mkmk."
thanks: "Code and data supporting this analysis are available at: https://github.com/CSCmaster/Final-Project"
author: "Jingyi Shen"
date: today
date-format: long
format: pdf
toc: true
toc-depth: 2
number-sections: true
bibliography: references.bib
link-citations: true
---

@@

### Data Summary Measures 
```{r}
#| label: tbl-data-summary
#| tbl-cap: Summary of Numerical Variables Across Census and Survey Data
#| echo: false
numeric_summary_df <- data.frame(
  Variable = c("Age in Census Data", "Age in Survey Data", 
               "Liberal Votes in Survey", "Conservative Votes in Survey"),
  Minimum = c(15.0, 18.0, 0.000, 0.000),
  Q1 = c(37.0, 36.0, 0.000, 0.000),
  Median = c(54.0, 53.0, 0.000, 0.000),
  Average = c(52.18, 51.3, 0.267, 0.249),
  Q3 = c(67.0, 66.0, 1.000, 0.000),
  Maximum = c(80.0, 97.0, 1.000, 1.000)
)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cleaned_survey_data <- read_parquet(here("data", "analysis_data","cleaned_survey_data.parquet"))
cleaned_census_data <- read_parquet(here("data", "analysis_data","cleaned_census_data.parquet"))
combined_data <- rbind(
  data.frame(Age = cleaned_survey_data$age, Dataset = "Survey"),
  data.frame(Age = cleaned_census_data$age, Dataset = "Census")
)
ggplot(combined_data, aes(x = Age, fill = Dataset)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 4) +
  scale_fill_manual(values = c("Survey" = "blue", "Census" = "red")) +
  labs(
    title = "Plot.1 Age Distribution Comparison",
    x = "Age",
    y = "Frequency"
  ) +
  xlim(min(combined_data$Age, na.rm = TRUE), max(combined_data$Age, na.rm = TRUE)) +
  theme_minimal()

@@

The overlapping histogram illustrates the age distributions from census and survey datasets, with the census data in red displaying a more concentrated and unimodal age distribution peaking between 50 and 60 years and tapering off sharply at 80. In contrast, the survey data, shown in blue, reveals a broader spread of ages with a right-skewed tail extending to 97 years, indicating the inclusion of older respondents. Both distributions share a common median age range, highlighted by the combined purple hue in the overlap, but the census data consistently show higher frequencies across most age bins. This visual comparison underscores the differing age profiles captured by the two datasets and a significant density of middle-aged respondents in both populations.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
combined_provinces <- rbind(
  data.frame(province = cleaned_survey_data$province, dataset = "Survey Dataset"),
  data.frame(province = cleaned_census_data$province, dataset = "Census Dataset")
)
ggplot(combined_provinces, aes(x = province, fill = dataset)) +
  geom_bar(position = "dodge") +
  labs(title = "Plot.2 Comparison of Province Variable", x = "Province", y = "Count") +  
  scale_fill_manual(values = c("Survey Dataset" = "red", "Census Dataset" = "blue")) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

The bar chart depicts the frequency distribution of respondents across different provinces as collected in two datasets: the census and survey data. Each province is represented on the x-axis, and the corresponding count of respondents is displayed on the y-axis. The bars are color-coded, with blue indicating the survey dataset and red representing the census dataset. The bars are placed side by side, or 'dodged', allowing for a direct visual comparison between the two datasets within each provincial category.

From a statistical perspective, Ontario has the highest frequency in the survey dataset, which substantially exceeds the count in the census dataset, indicating a larger sample or population size recorded in the survey for this province. Conversely, the counts for British Columbia and Alberta show a relatively balanced representation between the two datasets. The chart facilitates a comparative analysis of the datasets, revealing disparities and similarities in the population coverage across provinces, which could be indicative of sampling strategies or the demographic spread within the datasets.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
combined_data_education <- na.omit(rbind(
  data.frame(Education = cleaned_survey_data$education, Dataset = "Survey Dataset"),
  data.frame(Education = cleaned_census_data$education, Dataset = "Census Dataset")
))
education_summary <- combined_data_education %>%
  group_by(Education, Dataset) %>%
  summarise(Count = n(), .groups = 'drop')

@@

ggplot(education_summary, aes(x = Education, y = Count, group = Dataset, color = Dataset)) +
  geom_line() + 
  geom_point() + 
  scale_color_manual(values = c("Survey Dataset" = "red", "Census Dataset" = "blue")) +
  labs(title = "Line Chart Comparison of Education Levels",
       x = "Education Level",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

The line chart presented compares the distribution of education levels across two different datasets, one from a survey and another from a census, differentiated by colors—red for the survey and blue for the census. Education levels are divided into three categories along the x-axis: "Highly Educated", "Limited Education", and "Some Education", with the y-axis displaying the count of individuals within each category. A noticeable trend is the inverse relationship between the two datasets from "Highly Educated" to "Limited Education": the count for the Census dataset declines while that for the Survey dataset rises, suggesting a disparity in the representation of education levels between the two data sources.

This discrepancy highlights potential differences in sampling methods, response rates, or the representativeness of each dataset relative to the underlying population. The most pronounced difference is at the "Limited Education" level, where the Survey dataset shows significantly lower counts compared to the Census dataset, indicating a possible overrepresentation of individuals with limited education in the censaus or an underrepresentation in the survey. The data suggest that demographic, socioeconomic, or regional factors may influence the education level distribution in each dataset, and further statistical analysis could be warranted to determine the significance and cause of these differences. The differences observed in the chart might also reflect methodological variations in data collection or inherent biases within the sampling frames used.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

combined_data_religion <- rbind(
  data.frame(Religion = cleaned_survey_data$religion, Dataset = "Survey Dataset"),
  data.frame(Religion = cleaned_census_data$religion, Dataset = "Census Dataset")
)

@@

ggplot(combined_data_religion, aes(x = Religion, fill = Dataset)) +
  geom_bar(position = "dodge") +  # Position bars side by side
  labs(title = "Plot.4 Comparison of Religion", x = "Religion", y = "Count") +  
  scale_fill_manual(values = c("Survey Dataset" = "pink", "Census Dataset" = "red")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Optional: to improve readability of x-axis labels

The bar graph depicts the frequency of individuals categorized by religious affiliation within two datasets: census and survey. Each dataset's frequency is represented by a bar in distinct shades, with a darker red indicating the census dataset and a lighter pink representing the survey dataset. The x-axis bifurcates the data into two groups: those without religious affiliation ('NO') and those with religious affiliation ('YES').

The y-axis provides a count of individuals in each category. Notably, the census dataset shows a substantially higher count of individuals with a religious affiliation compared to those without. This pattern is less pronounced in the survey dataset, where the frequencies of individuals with and without religious affiliation are more comparable. The visual disparity in bar heights, especially within the 'YES' category, suggests a higher prevalence of religious affiliation reported in the census dataset than in the survey dataset.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
combined_data_income <- na.omit(rbind(
  data.frame(IncomeBeforeTax = cleaned_survey_data$income_before_tax, Dataset = "Survey Dataset"),
  data.frame(IncomeBeforeTax = cleaned_census_data$income_before_tax, Dataset = "Census Dataset")
))
ggplot(combined_data_income, aes(x = IncomeBeforeTax, fill = Dataset)) +
  geom_bar(position = "dodge") +
  labs(title = "Plot.5 Comparison of Income Before Tax", x = "Income Before Tax", y = "Count") +
  scale_fill_manual(values = c("Survey Dataset" = "orange", "Census Dataset" = "brown")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

The bar chart displays the distribution of income categories before tax across two datasets: census and survey. The income categories are defined as 'Lower Middle Class to Poor', 'Middle Class', and 'Upper Middle Class to Wealthy', and are represented on the x-axis. The y-axis measures the count of individuals within each income bracket.

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_long <- pivot_longer(na.omit(cleaned_survey_data), cols = starts_with("vote_"), names_to = "Party", values_to = "Vote")

ggplot(df_long, aes(x = Party, fill = as.factor(Vote))) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Plot.6 Comparison of Voting Preferences", x = "Party", y = "Count") +
  scale_fill_manual(values = c("red", "blue"), name = "Party Choice", labels = c("Did Not Vote", "Voted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

\begin{align}
y_i | \pi_i &\sim \text{Bernoulli}(\pi_i) \\
\text{logit}(\pi_i) &= \beta_0 + \alpha^{age}_{a[i]} + \alpha^{gender}_{g[i]} + \alpha^{province}_{p[i]} + \alpha^{religion}_{r[i]} + \alpha^{income}_{m[i]} + \alpha^{education}_{e[i]} \\
\beta_0 &\sim \text{Normal}(0, 2.5) \\
\alpha_{a}^{age} &\sim \text{Normal}(0, \sigma^{2}_{age}) \quad \text{for } a \text{ in } 1, 2, \ldots, A \\
\alpha_{g}^{gender} &\sim \text{Normal}(0, \sigma^{2}_{gender}) \quad \text{for } g \text{ in } Male,Female \\
\alpha_{p}^{province} &\sim \text{Normal}(0, \sigma^{2}_{province}) \quad \text{for } p \text{ in } BC, AB, MB, NB, NL, NS, ON, PE, QC, SK \\
\alpha_{r}^{religion} &\sim \text{Normal}(0, \sigma^{2}_{religion}) \quad \text{for } r \text{ in } YES,NO \\
\alpha_{m}^{income} &\sim \text{Normal}(0, \sigma^{2}_{income}) \quad \text{for } m \text{ in } Lower Class, Middle Class, Upper Class\\
\alpha_{e}^{education} &\sim \text{Normal}(0, \sigma^{2}_{education}) \quad \text{for } e \text{ in } Low Education, Some Education, High Education \\
\sigma^{2}_{age} &\sim \text{Exponential}(1) \\
\sigma^{2}_{gender} &\sim \text{Exponential}(1) \\
\sigma^{2}_{province} &\sim \text{Exponential}(1) \\
\sigma^{2}_{religion} &\sim \text{Exponential}(1) \\
\sigma^{2}_{income} &\sim \text{Exponential}(1) \\
\sigma^{2}_{education} &\sim \text{Exponential}(1) 
\end{align}

## Model justification

@@

Table2: Summary the residuals for Conservative Party:
```{r echo=FALSE, warning=FALSE}
liberal_residuals <- posterior_predict(liberal_model) - cleaned_survey_data$vote_liberal

@@

print(table_summary)
```
Table3: Summary the coefficient for Liberal Party:
```{r echo=FALSE, warning=FALSE}

@@

@@

```{r echo=FALSE, warning=FALSE}
pp_check(liberal_model, type = "hist")
```
Histogram of posterior predictive distributions from the liberal model. This plot compares the observed counts of the number of votes for the Liberal party (indicated by the dark line) with the simulated counts from the posterior predictive distribution (shown as histograms). The overlap between the observed counts and the predictive distribution suggests how well the model captures the data. Areas where the observed counts fall outside the bulk of the simulated distributions may indicate model misfit.

```{r echo=FALSE, warning=FALSE}
pp_check(conservative_model, type = "hist")
```

```{r echo=FALSE, warning=FALSE,  message=FALSE, fig.width=8, fig.height=6}
survey_data <- read_csv(here("data", "analysis_data", "cleaned_survey_data.csv"))
prior_covariates <- normal(location = 0, scale = 2.5, autoscale = FALSE)
prior_intercept <- normal(0, 2.5)
pp_plot <- posterior_vs_prior(liberal_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 10),
    axis.title.y = element_text(size = 7, face = "bold"),
    axis.title.x = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 7)
  ) +
  coord_flip() +
  ggtitle("Comparison of Posterior and Prior Distributions",
          subtitle = "This plot visualizes how posterior distributions of model parameters compare with their specified priors for the liberal model.") +
  xlab("Density") +
  ylab("Parameters")

# Print the plot
print(pp_plot)
```

```{r echo=FALSE, warning=FALSE,  message=FALSE, fig.width=8, fig.height=6}
prior_covariates <- normal(location = 0, scale = 2.5, autoscale = FALSE)
prior_intercept <- normal(0, 2.5)
pp_plot <- posterior_vs_prior(conservative_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 10),
    axis.title.y = element_text(size = 7, face = "bold"),
    axis.title.x = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 7)
  ) +
  coord_flip() +
  ggtitle("Comparison of Posterior and Prior Distributions",
          subtitle = "This plot visualizes how posterior distributions of model parameters compare with their specified priors for the conservative model.") +
  xlab("Density") +
  ylab("Parameters")

# Print the plot
print(pp_plot)

@@

>>>>

---
title: "Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election "
title-block-banner: true
abstract: "ojjojojojljk,mkmk."
thanks: "Code and data supporting this analysis are available at: https://github.com/CSCmaster/Final-Project"
author: "Jingyi Shen"
date: today
date-format: long
format: pdf
toc: true
toc-depth: 2
number-sections: true
bibliography: references.bib
link-citations: true
editor: 
  markdown: 
    wrap: sentence
---

@@

### Data Summary Measures

```{r}
#| label: tbl-data-summary
#| tbl-cap: Summary of Numerical Variables Across Census and Survey Data
#| echo: false
numeric_summary_df <- data.frame(
  Variable = c("Age in Census Data", "Age in Survey Data", 
               "Liberal Votes in Survey", "Conservative Votes in Survey"),
  Minimum = c(15.0, 18.0, 0.000, 0.000),
  Q1 = c(37.0, 36.0, 0.000, 0.000),
  Median = c(54.0, 53.0, 0.000, 0.000),
  Average = c(52.18, 51.3, 0.267, 0.249),
  Q3 = c(67.0, 66.0, 1.000, 0.000),
  Maximum = c(80.0, 97.0, 1.000, 1.000)
)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-age-distribution
#| tbl-cap: "Age Distribution Comparison Between Survey and Census Data"
#| echo: false
cleaned_survey_data <- read_parquet(here("data", "analysis_data","cleaned_survey_data.parquet"))
cleaned_census_data <- read_parquet(here("data", "analysis_data","cleaned_census_data.parquet"))
combined_data <- rbind(
  data.frame(Age = cleaned_survey_data$age, Dataset = "Survey"),
  data.frame(Age = cleaned_census_data$age, Dataset = "Census")
)
ggplot(combined_data, aes(x = Age, fill = Dataset)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 4) +
  scale_fill_manual(values = c("Survey" = "blue", "Census" = "red")) +
  labs(
    x = "Age",
    y = "Frequency"
  ) +
  xlim(min(combined_data$Age, na.rm = TRUE), max(combined_data$Age, na.rm = TRUE)) +
  theme_minimal()

@@

@fig-age-distribution illustrates the age distributions from census and survey datasets, with the census data in red displaying a more concentrated and unimodal age distribution peaking between 50 and 60 years and tapering off sharply at 80.
In contrast, the survey data, shown in blue, reveals a broader spread of ages with a right-skewed tail extending to 97 years, indicating the inclusion of older respondents.
Both distributions share a common median age range, highlighted by the combined purple hue in the overlap, but the census data consistently show higher frequencies across most age bins.
This visual comparison underscores the differing age profiles captured by the two datasets and a significant density of middle-aged respondents in both populations.
\newpage

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-province-distribution
#| fig-cap: "Comparison of Province Variable Between Survey and Census Data"
#| echo: false
combined_provinces <- rbind(
  data.frame(province = cleaned_survey_data$province, dataset = "Survey Dataset"),
  data.frame(province = cleaned_census_data$province, dataset = "Census Dataset")
)
ggplot(combined_provinces, aes(x = province, fill = dataset)) +
  geom_bar(position = "dodge") +
  labs(x = "Province", y = "Count") +  
  scale_fill_manual(values = c("Survey Dataset" = "red", "Census Dataset" = "blue")) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

@fig-province-distribution depicts the frequency distribution of respondents across different provinces as collected in two datasets: the census and survey data.
Each province is represented on the x-axis, and the corresponding count of respondents is displayed on the y-axis.
The bars are color-coded, with blue indicating the survey dataset and red representing the census dataset.
The bars are placed side by side, or 'dodged', allowing for a direct visual comparison between the two datasets within each provincial category.

From a statistical perspective, Ontario has the highest frequency in the survey dataset, which substantially exceeds the count in the census dataset, indicating a larger sample or population size recorded in the survey for this province.
Conversely, the counts for British Columbia and Alberta show a relatively balanced representation between the two datasets.
The chart facilitates a comparative analysis of the datasets, revealing disparities and similarities in the population coverage across provinces, which could be indicative of sampling strategies or the demographic spread within the datasets.
\newpage

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-education-distribution
#| fig-cap: "Line Chart Comparison of Education Levels Between Survey and Census Data"
#| echo: false
combined_data_education <- na.omit(rbind(
  data.frame(Education = cleaned_survey_data$education, Dataset = "Survey Dataset"),
  data.frame(Education = cleaned_census_data$education, Dataset = "Census Dataset")
))
education_summary <- combined_data_education %>%
  group_by(Education, Dataset) %>%
  summarise(Count = n(), .groups = 'drop')

@@

ggplot(education_summary, aes(x = Education, y = Count, group = Dataset, color = Dataset)) +
  geom_line() + 
  geom_point() + 
  scale_color_manual(values = c("Survey Dataset" = "red", "Census Dataset" = "blue")) +
  labs(
       x = "Education Level",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

@fig-education-distribution presented compares the distribution of education levels across two different datasets, one from a survey and another from a census, differentiated by colors—red for the survey and blue for the census.
Education levels are divided into three categories along the x-axis: "Highly Educated", "Limited Education", and "Some Education", with the y-axis displaying the count of individuals within each category.
A noticeable trend is the inverse relationship between the two datasets from "Highly Educated" to "Limited Education": the count for the Census dataset declines while that for the Survey dataset rises, suggesting a disparity in the representation of education levels between the two data sources.

This discrepancy highlights potential differences in sampling methods, response rates, or the representativeness of each dataset relative to the underlying population.
The most pronounced difference is at the "Limited Education" level, where the Survey dataset shows significantly lower counts compared to the Census dataset, indicating a possible overrepresentation of individuals with limited education in the censaus or an underrepresentation in the survey.
The data suggest that demographic, socioeconomic, or regional factors may influence the education level distribution in each dataset, and further statistical analysis could be warranted to determine the significance and cause of these differences.
The differences observed in the chart might also reflect methodological variations in data collection or inherent biases within the sampling frames used.
\newpage

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-religion-distribution
#| fig-cap: "Bar Chart Comparison of Religion Across Survey and Census Data"
#| echo: false
combined_data_religion <- rbind(
  data.frame(Religion = cleaned_survey_data$religion, Dataset = "Survey Dataset"),
  data.frame(Religion = cleaned_census_data$religion, Dataset = "Census Dataset")
)

@@

ggplot(combined_data_religion, aes(x = Religion, fill = Dataset)) +
  geom_bar(position = "dodge") +  
  labs(x = "Religion", y = "Count") +  
  scale_fill_manual(values = c("Survey Dataset" = "pink", "Census Dataset" = "red")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  
```

@fig-religion-distribution depicts the frequency of individuals categorized by religious affiliation within two datasets: census and survey.
Each dataset's frequency is represented by a bar in distinct shades, with a darker red indicating the census dataset and a lighter pink representing the survey dataset.
The x-axis bifurcates the data into two groups: those without religious affiliation ('NO') and those with religious affiliation ('YES').

The y-axis provides a count of individuals in each category.
Notably, the census dataset shows a substantially higher count of individuals with a religious affiliation compared to those without.
This pattern is less pronounced in the survey dataset, where the frequencies of individuals with and without religious affiliation are more comparable.
The visual disparity in bar heights, especially within the 'YES' category, suggests a higher prevalence of religious affiliation reported in the census dataset than in the survey dataset.
\newpage

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-income-distribution
#| fig-cap: "Bar Chart Comparison of Income Before Tax Between Survey and Census Data"
#| echo: false
combined_data_income <- na.omit(rbind(
  data.frame(IncomeBeforeTax = cleaned_survey_data$income_before_tax, Dataset = "Survey Dataset"),
  data.frame(IncomeBeforeTax = cleaned_census_data$income_before_tax, Dataset = "Census Dataset")
))
ggplot(combined_data_income, aes(x = IncomeBeforeTax, fill = Dataset)) +
  geom_bar(position = "dodge") +
  labs(x = "Income Before Tax", y = "Count") +
  scale_fill_manual(values = c("Survey Dataset" = "orange", "Census Dataset" = "brown")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

@fig-income-distribution displays the distribution of income categories before tax across two datasets: census and survey.
The income categories are defined as 'Lower Middle Class to Poor', 'Middle Class', and 'Upper Middle Class to Wealthy', and are represented on the x-axis.
The y-axis measures the count of individuals within each income bracket.

@@

\newpage

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-voting-preferences
#| fig-cap: "Bar Chart Comparison of Voting Preferences"
#| echo: false
df_long <- pivot_longer(na.omit(cleaned_survey_data), cols = starts_with("vote_"), names_to = "Party", values_to = "Vote")

ggplot(df_long, aes(x = Party, fill = as.factor(Vote))) +
  geom_bar(position = "dodge", stat = "count") +
  labs(x = "Party", y = "Count") +
  scale_fill_manual(values = c("red", "blue"), name = "Party Choice", labels = c("Did Not Vote", "Voted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@fig-voting-preferences contrasts the voting preferences for the Conservative and Liberal parties, delineating counts of participants who voted (blue bars) against those who did not (red bars).
Notably, there is a substantial number of non-voters in the Conservative category, exceeding their Liberal counterparts, suggesting a larger disenchantment or mobilization challenge within the Conservative base.
Despite this, the voter turnout (blue bars) is fairly balanced between the parties, indicating a competitive engagement among those who participated in the vote.
The overall higher presence of red bars signifies a general trend of voter disengagement, which may warrant further investigation through regression analysis and post-stratification to grasp the full scope of electoral dynamics and the broader implications for the Canadian democratic process.
\newpage

@@

```{=tex}
\begin{align}
y_i | \pi_i &\sim \text{Bernoulli}(\pi_i) \\
\text{logit}(\pi_i) &= \beta_0 + \alpha^{age}_{a[i]} + \alpha^{gender}_{g[i]} + \alpha^{province}_{p[i]} + \alpha^{religion}_{r[i]} + \alpha^{income}_{m[i]} + \alpha^{education}_{e[i]} \\
\beta_0 &\sim \text{Normal}(0, 2.5) \\
\alpha_{a}^{age} &\sim \text{Normal}(0, \sigma^{2}_{age}) \quad \text{for } a \text{ in } 1, 2, \ldots, A \\
\alpha_{g}^{gender} &\sim \text{Normal}(0, \sigma^{2}_{gender}) \quad \text{for } g \text{ in } Male,Female \\
\alpha_{p}^{province} &\sim \text{Normal}(0, \sigma^{2}_{province}) \quad \text{for } p \text{ in } BC, AB, MB, NB, NL, NS, ON, PE, QC, SK \\
\alpha_{r}^{religion} &\sim \text{Normal}(0, \sigma^{2}_{religion}) \quad \text{for } r \text{ in } YES,NO \\
\alpha_{m}^{income} &\sim \text{Normal}(0, \sigma^{2}_{income}) \quad \text{for } m \text{ in } Lower Class, Middle Class, Upper Class\\
\alpha_{e}^{education} &\sim \text{Normal}(0, \sigma^{2}_{education}) \quad \text{for } e \text{ in } Low Education, Some Education, High Education \\
\sigma^{2}_{age} &\sim \text{Exponential}(1) \\
\sigma^{2}_{gender} &\sim \text{Exponential}(1) \\
\sigma^{2}_{province} &\sim \text{Exponential}(1) \\
\sigma^{2}_{religion} &\sim \text{Exponential}(1) \\
\sigma^{2}_{income} &\sim \text{Exponential}(1) \\
\sigma^{2}_{education} &\sim \text{Exponential}(1) 
\end{align}
```
## Model justification

@@

Table2: Summary the residuals for Conservative Party:

```{r echo=FALSE, warning=FALSE}
liberal_residuals <- posterior_predict(liberal_model) - cleaned_survey_data$vote_liberal

@@

print(table_summary)
```

Table3: Summary the coefficient for Liberal Party:

```{r echo=FALSE, warning=FALSE}

@@

\newpage

@@

```{r fig-liberal-model-1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Posterior Predictive Check: Observed vs. Simulated Vote Counts in the Liberal Model"}
library(bayesplot)

liberal_model_plot <- pp_check(liberal_model, type = "hist") +
  theme(plot.title = element_text(size = 10))

print(liberal_model_plot)
```

@fig-liberal-model-1 illustrates a posterior predictive check for the liberal model, comparing observed vote counts with those generated by the model's simulations.
The dark solid line represents the observed data, while the lighter blue lines represent simulated vote counts from the posterior distribution.
The close tracking of the simulated lines to the observed line suggests that the model can effectively replicate the patterns seen in the real vote counts.
This indicates a good model fit, as the simulations encompass the range of observed data, providing confidence in the model's predictions.
The alignment of the simulations with the observed data points to the adequacy of the model's structure and the priors in capturing the underlying trends of the voting behavior it aims to represent.
The graph also visualizes the variance in the simulations, reflecting the model's probabilistic nature and the inherent uncertainty in predicting complex phenomena such as voting behavior.

Histogram of posterior predictive distributions from the liberal model.
This plot compares the observed counts of the number of votes for the Liberal party (indicated by the dark line) with the simulated counts from the posterior predictive distribution (shown as histograms).
The overlap between the observed counts and the predictive distribution suggests how well the model captures the data.
Areas where the observed counts fall outside the bulk of the simulated distributions may indicate model misfit.
\newpage

```{r fig-conservative-model, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Posterior Predictive Check: Observed vs. Simulated Vote Counts in the Conservative Model"}
library(bayesplot)

conservative_model_plot <- pp_check(conservative_model, type = "hist") + 
  theme(plot.title = element_text(size = 14)) 
print(conservative_model_plot)

@@

@fig-conservative-model performs a posterior predictive check for a conservative model by overlaying observed vote counts against simulated data derived from the model's posterior distribution.
The observed data is denoted by a solid line, while the simulations from the posterior are indicated by lighter lines.
The congruence between these two sets of lines—where the simulated data traces the shape and trend of the observed data closely—suggests that the model has a high level of accuracy in reflecting the observed outcomes.
This indicates a successful fit of the model to the data, affirming the appropriateness of the selected priors and the model's structure.
The spread of the lighter lines around the darker one provides insight into the inherent uncertainty of the model's predictions, which is an expected characteristic of predictive modelling in complex, real-world situations.

\newpage

```{r fig-liberal-model, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=6, fig.cap="Comparison of Posterior and Prior Distributions in the Liberal Model"}
library(readr)
library(bayesplot)
library(here)

survey_data <- read_csv(here("data", "analysis_data", "cleaned_survey_data.csv"))

prior_covariates <- normal(location = 0, scale = 2.5, autoscale = FALSE)
prior_intercept <- normal(0, 2.5)

pp_plot <- posterior_vs_prior(liberal_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 10),
    axis.title.y = element_text(size = 7, face = "bold"),
    axis.title.x = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 7)
  ) +
  coord_flip() +
  xlab("Density") +
  ylab("Parameters")

print(pp_plot)

@fig-liberal-model portrays the posterior and prior distributions for parameters in a Bayesian liberal model, with a stark divergence in the intercept suggesting significant data-driven adjustments from the model's baseline estimation.
Parameters representing demographic and geographical factors are plotted as colored dots along an axis, indicating median values and the extent of uncertainty.
While the priors and posteriors largely overlap, indicating consistent estimates for many parameters, the intercept's noticeable shift underscores the data's impactful role in refining the model, particularly highlighting the updated understanding of the model's fundamental starting point.

\newpage

```{r fig-conservative-posterior, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=6, fig.cap="Comparison of Posterior and Prior Distributions in the Conservative Model"}
library(bayesplot)

prior_covariates <- normal(location = 0, scale = 2.5, autoscale = FALSE)
prior_intercept <- normal(0, 2.5)

pp_plot <- posterior_vs_prior(conservative_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 10),
    axis.title.y = element_text(size = 7, face = "bold"),
    axis.title.x = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 7)
  ) +
  coord_flip() +
  xlab("Density") +
  ylab("Parameters")

print(pp_plot)

@fig-conservative-posterior presents a side-by-side comparison of the median and spread of prior and posterior distributions for the parameters in a Bayesian model, applied to the Conservative context.
Key findings include the posterior distribution of the "age" parameter mirroring its prior, which validates the initial assumption about its influence.
In contrast, the intercept and sigma parameters show slight shifts, indicating the data's role in refining these estimates.
The placement of medians for other parameters—like sex, religion, and province—demonstrates the data's impact in confirming or adjusting prior beliefs.
Overall, the alignment of posterior distributions with priors confirms the model's accuracy and the informative value of the data in capturing the effects on the dependent variable, suggesting a robust model fit.

\newpage
